---
title: "Bechdel"
author: "Anna Smyrnova"
date: "August 21, 2019"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
#pacman::p_load(tidyverse, dplyr, ggplot2, gridExtra, VIM, sophisthse, car, MASS, corrplot)

library(dplyr)          # data manipulation
library(tidyverse)      # data manipulation, exploration and visualization
library(pander)         # An R 'Pandoc' Writer
library(corrplot)
library(MASS)
library(car)
library(sophisthse)     # labels
library(VIM)            # Visualization and Imputation of Missing Values
library(caret)

library(ggplot2)        # plotting & data
theme_set(
  theme_bw() +
    theme(legend.position = "top")
  )
library(gridExtra)      # provides side-by-side plotting

knitr::opts_chunk$set(echo = TRUE)
require("knitr")
opts_knit$set(root.dir = "~/R/PA_coding_challenge_AS")
```

#Part 1: Data setup
You have a dataset called "Bechdel.xlsx" in Excel format. Import it into R and prepare the dataset for analysis.

```{r reading dataset, message=FALSE, warning=FALSE}
library(readxl)
bechdel <- read_excel("bechdel.xlsx", col_types = c("numeric", 
     "text", "numeric", "numeric", "numeric", 
     "text", "numeric", "numeric", "numeric", 
     "text", "skip", "skip"), skip = 1)

bechdel$year<-as.integer(bechdel$year)
head(bechdel)
str(bechdel)
```
### Labels

```{r labels, message=FALSE, warning=FALSE}
label <- read_excel("bechdel.xlsx", sheet = "bechdel", 
     range = "A1:J1", col_names = FALSE)
## to characters
label = as.character(label[1,])

## assign labels
set_variable_labels(bechdel, label)
rm(label)
```

### Missing values

```{r missing values, fig.cap="Missing data proportions and patterns"}
#The visuals were created using the VIM package. 
aggr(bechdel, cex.axis = 0.7)
```  
```{r summary}
pander(summary(bechdel))
```
Looks like there are no outliers/errors, all data looks quite reasonable (for example, Avatar had the budget of USD425,000,000	 and worldwide gross of USD2,789,705,275 in 2009).
The distribution of missing data indicates that 4 variables have missing data points, but proportion of missing values are very small, imputation will not be necessary, a complete case analysis will suffice.

```{r complete case}
bechdel<-bechdel[complete.cases(bechdel), ]
```

# Part 2: Data analysis

i) Summarise the data using any descriptive statistics, inferential statistics, or visualisations that you deem appropriate. Address the following questions as well:
– What percentage of movies passed the test, based on the binary definition?

```{r }
# What percentage of movies passed the test, based on the binary definition?
bechdel %>%
  group_by(binary) %>%
  summarise (number = n() ) %>%
  mutate(freq = number / sum(number)) %>% ungroup()
```

- How does the percentage of movies passing change over time?

```{r freq PASS by year}
pct_year<- bechdel %>%
  group_by(year, binary) %>%
  summarise (number = n() ) %>%
  mutate(freq = number / sum(number)) %>% ungroup()

total_year<-bechdel %>%
  group_by(year)%>%
  summarise (total = n() ) %>% ungroup 

pct_pass <- pct_year %>%
  dplyr::select(year, binary, number, freq) %>%
  full_join(total_year, by="year")

rm(pct_year, total_year)
```


```{r plot pct PASS by year}
figure1 <- pct_pass %>% dplyr::filter(binary=="PASS")

ggplot(figure1) +
geom_area( fill="lightblue", alpha=.5, aes(y = total, x=year)) +
geom_line( color="red", alpha=1, aes(y = freq*100, x=year)) +
ggtitle("Total number of films and proportion of movies passed the Bechdel test") +
scale_y_continuous(sec.axis = sec_axis(~., name = "Total number of films")) +
labs(y = "Films that pass the test [%] ", x = "Year") +
theme_bw()+
scale_x_continuous(breaks=seq(1970, 2013, 10))+
theme(legend.text = element_text(colour="blue", size=10, face="bold"))

```


As we use more data, the proportion of movies passed the Bechdel test tends to show less variation.

```{r additional categorical variables}
#decade categories
bechdel$decade<-cut(bechdel$year, c(1969, 1980,1990,2000, 2014), right=FALSE)

bechdel$decade<-factor(bechdel$decade, labels = c("1970x", "1980x", "1990x", "2000x"))

decade_ds <- bechdel %>%
  group_by(decade, binary) %>%
  summarise (number = n() ) %>%
  mutate(freq = number / sum(number)) %>% ungroup()

pander(decade_ds)
```

```{r tendency by decade}
# the overall tendency by decade
pct_decade<-  bechdel %>%
  group_by(decade, binary) %>%
  summarise (number = n() ) %>%
  mutate(freq = round(number / sum(number),2)) %>% ungroup()


ggplot(data=pct_decade, aes(x=decade, y=freq, fill=binary)) +
  geom_bar(stat="identity",position=position_dodge())+
geom_text(aes(label=freq), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5)+
  scale_fill_brewer(palette="Paired")+
ggtitle("Changes in proportion of movies passed the Bechdel test during 1970-2013, \nby decade")+
  theme(legend.text = element_text(colour="grey", size=10, face="bold"))

```



```{r Box Office }
budget_year<-  bechdel %>%
  group_by(year, binary) %>%
  summarise (mean_budget=mean(budget_2013), mean_domgross=mean(domgross_2013), mean_intgross=mean(intgross_2013)) %>%
  mutate(roi=(mean_domgross+mean_intgross)/mean_budget, mean_total=mean_domgross+mean_intgross)
```

```{r tendency gross}
p1<-ggplot(data=budget_year, aes(x=year, y=mean_budget, color=binary))+
  geom_line()+
  scale_color_brewer(palette="Paired")+
ggtitle("Mean budget for Pass and Fail \nmovies during 1970-2013")+
  theme(legend.text = element_text(colour="grey", size=10, face="bold"))

p2<-ggplot(data=budget_year, aes(x=year, y=mean_total, color=binary))+
  geom_line()+
  scale_color_brewer(palette="Paired")+
ggtitle("Mean gross (domestic + inter) for \nPass and Fail movies during 1970-2013")+
  theme(legend.text = element_text(colour="grey", size=10, face="bold"))

p3<-ggplot(data=budget_year, aes(x=year, y=roi, color=binary))+
  geom_line()+
  scale_color_brewer(palette="Paired")+
ggtitle(" ROI for Pass and Fail movies \nduring 1970-2013")+
  theme(legend.text = element_text(colour="grey", size=10, face="bold"))

p4<-ggplot(data=budget_year, aes(x=year, y=mean_budget, color=binary))+
  geom_line()+
  scale_color_brewer(palette="Paired")+
ggtitle("Mean budget for Pass and Fail \nmovies during 1970-2013")+
  geom_smooth(method = "loess") +
  theme(legend.text = element_text(colour="grey", size=10, face="bold"))

p5<-ggplot(data=budget_year, aes(x=year, y=mean_total, color=binary))+
  geom_line()+
  scale_color_brewer(palette="Paired")+
ggtitle("Mean gross (domestic + inter)for \nPass and Fail movies during 1970-2013")+
  geom_smooth(method = "loess") +
  theme(legend.text = element_text(colour="grey", size=10, face="bold"))

p6<-ggplot(data=budget_year, aes(x=year, y=roi, color=binary))+
  geom_line()+
  scale_color_brewer(palette="Paired")+
ggtitle(" ROI for Pass and Fail movies \nduring 1970-2013")+
  geom_smooth(method = "loess") +
  theme(legend.text = element_text(colour="grey", size=10, face="bold"))

grid.arrange(p1, p4,  ncol = 2)
grid.arrange(p2, p5,  ncol = 2)
grid.arrange(p3, p6,  ncol = 2)

```

```{r mean gross}
mean_gross <-aggregate(budget_year[, 3:7], list(budget_year$binary), mean)
ratio<-mean_gross[1, 2:6]/mean_gross[2, 2:6]

pander(mean_gross)
#ratio FAIL to PASS
pander(ratio)

```

ii) Propose a model that will predict whether a movie passes or fails (the binary variable) based
on whatever other variables in the data are available that you think are useful.

We can start with logistic regression. Check the correlation between the variables:
```{r}
mydata <- bechdel %>%
  dplyr::select_if(is.numeric) 
predictors <- colnames(mydata)
bechdel.cor = cor(bechdel[predictors])
corrplot(bechdel.cor, type = 'upper', diag = T, addCoef.col = "white")
```


```{r}
#list of variables
dput(names(bechdel))

#bechdel[predictors] <- scale(bechdel[predictors])

#add binary output
bechdel$output <- ifelse(bechdel$binary=="PASS", 1, 0)
```

```{r glm.fit1}
glm.fit1 <- glm(output ~ year + budget + domgross + intgross +budget_2013 + domgross_2013 + intgross_2013, data = bechdel, family = "binomial")
#testing the importance of predictors
dropterm(glm.fit1, test = "F")
```

```{r glm.fit2}
# drop year and budget
glm.fit2 <- glm(output ~  domgross + intgross +budget_2013 + domgross_2013 + intgross_2013, data = bechdel, family = "binomial")
summary(glm.fit2)
# checking VIF
vif(glm.fit2)
```

VIF suggests presence of very strong multicollinearity, so we have to deal with it. 
Examine the behaviour of predictors:

```{r}

probabilities <- predict(glm.fit2, type = "response")
mydata <- mydata %>%
  mutate(logit = log(probabilities/(1-probabilities))) %>%
  tidyr::gather(key = "predictors", value = "predictor.value", -logit)

ggplot(mydata, aes(logit, predictor.value))+
  geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth(method = "loess") + 
  theme_bw() + 
  facet_wrap(~predictors, scales = "free_y")
```

```{r}
test <- read.csv("test.csv")
test<-test[complete.cases(test), ]
str(test)
```

```{r}
glm.probs <- predict(glm.fit2, 
                    newdata = test, 
                    type = "response")

glm.pred <- ifelse(glm.probs > 0.5, 1, 0)
test$prediction <- as.factor(glm.pred)
```


```{r}
# Model performance

test$binary_correct <- substr(test$code, 5,8)

test$correct <- as.factor(ifelse(test$binary_correct=="PASS", 1, 0))
accuracy <- table(test$prediction, test$correct)
sum(diag(accuracy))/sum(accuracy)
confusionMatrix(test$prediction, test$correct)
```

not so exciting results...

```{r}
#Simply save your predictions in a csv file with the variable 'mdb' to indicate the
#movie and your classification of 0 or 1.

test_prediction <- test %>% dplyr::select(imdb, prediction) %>% rename(mdb=imdb)
write.csv(test_prediction, "test_prediction.csv")
```



